\chapter{Introduction and Background}


\section{Summary}

\begin{summary}[Continuous functions on $ \Omega $ and $ \closure{\Omega} $ ]
	Let $ \Omega \subset \R^n $ be an open set. $ C(\Omega) $ denotes the set of all continuous functions defined on $ \Omega $. Similarly, $ C(\closure{\Omega}) $ denotes the set of all continuous functions defined on the closure of $ \Omega $.
	
	For any $ f \in C(\Omega) $ we have $ f \in C(\closure{\Omega}) $. However, for the converse, if $ g \in C(\closure{\Omega}) $, then if $ g $ is uniformally continuous and $ \Omega $ is bounded, then $ g $ can continuously be extended to $ \partial \Omega $. Note that $ C(\Omega) $ functions can behave badly near $ \partial \Omega $. For instance, consider the function $ f:(0,1) \to \R $ given by $ f(x) = \sin(1/x) $.
\end{summary}

\begin{summary}[The space of continuous $ 2\pi $ periodic functions]
	Consider the space of continouse functions defined on $ \R $, i.e. $ C(\R) $. An important subset of this set is $ C_p(2\pi) $ which is the set of all continuous $ 2\pi $ periodic functions where for $ f \in C_p(2\pi) $ we have
	\[ f(x+2\pi) = f(x), \qquad x\in \R. \]
	This set, is in one-to-one correspondence with the set of all continuous function defined from the manifold $ S^1 $, or equivalently $ \R/2\pi\Z $ to $ \R $.
\end{summary}

\begin{summary}[Basis for the set of polynomials]
	Let $ \mathbb{P}_n $ denote the set of all polynomials defined on $ \R $ with degree less than or equal to $ n $. Then a basis for this linear space is 
	\[ \mathbb{B} = \set{1,x,\cdots, x^n}. \]
	Thus the dimension of this space is $ n+1 $.
	
	Now, consider a \emph{linear subspace} of this space, the set of all polynomials that vanishes at $ 0 $ and $ 1 $ denoted by
	\[ \mathbb{P}_{n,0} = \set{p \in \mathbb{P}_n\ |\ p(0)= p(1) = 0}. \]
	A basis for this linear subspace can be given as
	\[ \mathbb{B}_{n,0} = \set{x(1-x),x^2(1-x),\cdots,x^{n-1}(1-x)}. \]
	Thus the dimension of this linear subspace is $ \dim(\mathbb{P}_n) - 2 $. The difference two in the dimension comes from the fact that polynomials in $ \mathbb{P}_{n,0} $ vanished at two points of their domain. Thus the set of all polynomials of degree $ n $ that vanish at $ n $ points of their domain is a 1 dimensional linear subspace of $ \mathbb{P}_n $.
\end{summary}


\begin{summary}[Normed space $ \R^d $]
	Consider the linear space $ \R^d $. Then the followings are the common norms for this space.
	\begin{align*}
		&\norm{x}_p = \left( \sum_{i=1}^d \abs{x_i}^p \right)^{1/p}, \qquad \text{for $ 1\leq p < \infty $}.\\
		&\norm{x}_\infty = \max_{1\leq i\leq d} \abs{x_i}.
	\end{align*}
	\begin{proposition}
		In $ \R^d $ we have for all $ x\in \R^d  $
		\[ \norm{x}_\infty \leq \norm{x}_p \leq d^{1/p}\norm{x}_\infty.  \]
	\end{proposition}
	\begin{remark}
		As a simple corollary of the proposition above we can see
		\[ \norm{x}_\infty = \lim_{p\to \infty} \norm{x}_p. \]
	\end{remark}
	\begin{proposition}[Equivalence of norms]
		On a finite dimensional space all norms are equivalent.
	\end{proposition}
	\begin{remark}
		The proposition above dose not hold true on infinite dimensional spaces. In those space, some norms has more stronger sense of convergence than others.
	\end{remark}
\end{summary}


\begin{summary}[Normed space $ C(\Omega) $ ]
	Let $ V = C[0,1] $ denote the linear space of all continuous function defined on $ [0,1] $. Define the following norms
	\begin{align*}
		& \norm{f}_p = \left(  \int_{0}^{1} \abs{f(x)}^p\ dx \right)^{1/p}, \qquad 1\leq p < \infty \\
		& \norm{f}_\infty = \sup_{0\leq x\leq 1} \abs{f(x)}
	\end{align*}
	The norm $ \norm{x}_\infty $ is a natural norm for this space that is also called \emph{uniform norm}.
	\begin{proposition}
		For the norms given above we have
		\[ \norm{v}_{p} \leq \norm{v}_\infty \qquad \forall v \in V. \]
		This implies that the convergence under the uniform norm $ \norm{\cdot}_\infty $ implies the convergence under the norm $ \norm{\cdot}_p $. Note that the converse is not true.
	\end{proposition}
	\begin{remark}
		A very good example to see the proposition above is $ f:[0,1]\to \R $ given by
		\[ f(x) = \begin{cases}
			1-nx, &\qquad 0\leq x\leq 1/n,\\
			0, &\qquad 1/n < x \leq 1.
		\end{cases} \]
	\end{remark}
	\begin{proposition}
		Let $ \Omega \subset \R^d $ and let $ \closure{\Omega} $ denote its closure. Then the space $ C(\closure{\Omega}) $ with the norm $ \norm{\cdot}_\infty $ is a Banach space. However, this space is not a Banach space with $ \norm{\cdot}_p $ for $ 1\leq p < \infty $.
	\end{proposition}
	\begin{remark}
		The proposition above is true since the continuoity of a sequence of functions persists under the uniform continuity. A good example for the second statement is the function $ f: [0,1] \to \R $
		\[ 
		f(x) = \begin{cases}
			0, &\qquad 0\leq x\leq 1/2 - 1/(2n),\\
			n(x-1/2+1/(2n)), &\qquad 1/2-1/(2n) \leq x \leq 1/2+1/(2n), \\
			1, &\qquad 1/2 + 1/(2n) \leq x \leq 1.
		\end{cases}
		 \]
	\end{remark}
\end{summary}

\begin{summary}[Normed space $ C^k(\Omega) $]
	Let $ \Omega \subset \R $. The space $ C^k(\closure{\Omega}) $ is the set of all $ k $ times continuously differentiable functions. Define the following metric on this space
	\begin{align*}
		&\norm{f}_{k,p} = \left( \sum_{i=1}^{k} \norm{f^{(i)}}_p^p \right)^{1/p}, \qquad 1\leq p < \infty. \\
		&\norm{f}_{k,\infty} = \max_{1\leq i \leq k} \norm{f^{(i)}}_p.
	\end{align*}
	The natural basis for this space is $ \norm{f}_{k,\infty} $.
	\begin{proposition}[$ C^k $ is a Banach space]
		The space $ C^k $ is complete under the norm $ \norm{\cdot}_{k,\infty} $.
	\end{proposition}
\end{summary}


\begin{summary}[Completion of $ C(\closure{\Omega}) $]
	The space $ C(\closure{\Omega}) $ is not complete under the norm $ \norm{\cdot}_{p} $ for $ 1\leq p < \infty $. Its completion is the space of \emph{Lebesgue integrable functions} $ L^p $.
\end{summary}
\begin{summary}[Completion of $ C^k(\closure{\Omega}) $]
	The space $ C^k(\closure{\Omega}) $ is not complete under the norm $ \norm{\cdot}_{k,p} $ for $ 1\leq p < \infty $. Its completion is the \emph{Sobolev spaces}.
\end{summary}

\begin{summary}[Norm changing the topology in action!]
	Consider the spaces $ V = C^1[0,1] \subset C[0,1] $, and $ W = C[0,1]$, and the linear operator 
	\[ T = \frac{d}{dx}: V \to W. \]
	Consider the same infinity norm $ \norm{\cdot}_{\infty} $ for both $ V $ and $ W $. Let $ \set{f_n} $ be a sequence of functions in $ V $ defined as
	\[ f_n(x) = \frac{1}{n} \sin(2^n\pi x). \]
	It is evident that $ \norm{f_n}_\infty \to 0 $ as $ n\to 0 $. However, $ \norm{f'_n}_\infty \to \infty $ as $ n\to \infty $. Geometrically, we can feel that the sequence $ \set{f_n} $ sort of moves towards the origin of the space $ (V,\norm{\cdot}_\infty) $ while $ \set{f'_n} $ shoots to infinity in the space $ (W, \norm{\cdot}_\infty) $. 
	
	However, if we change the norm of space $ V $ to the standard norm of $ C^1[0,1] $, i.e.
	\[ \norm{f}_\infty^1 = \max\set{\norm{f}_\infty,\norm{f'}_\infty}, \]
	then we can see that the sequence $ \set{f} $ is not moving towards the origin, but shoots of to the infinity of the space $ (V,\norm{\cdot}_\infty^1) $. From this example it is clear that how norm induces topology. A sequence that originally was moving towards the origin in one topology, shoots off to the infinity in another topology.
\end{summary}

\begin{summary}[Continuity of differentiation operator]
	According to the summary box above, the differentiation operator 
	\[ T_1 = \frac{d}{dx}: C^1[0,1]\subset (C[0,1],\norm{\cdot}_\infty) \to (C[0,1],\norm{\cdot}_\infty) \]
	is not continuous, but the operator
	\[ T_2 = \frac{d}{dx}: (C^1[0,1],\norm{\cdot}_\infty^1) \to (C[0,1],\norm{\cdot}_\infty) \]
	is continuous.
\end{summary}



\input{sections/introAndBackgroundSolvedProblems.tex}