\documentclass[10pt]{article}
%\usepackage[sc]{mathpazo}
\usepackage{commands}
\hypersetup{colorlinks = true}

\let\oldphi\phi 
\let\phi\varphi 
\let\varphi\oldphi

\definecolor{fom}{RGB}{0,153,139}

\usepackage [autostyle, english = american]{csquotes}



\begin{document}



\heading{Mathematical Proof}

This lecture note contains the material in the course MATH 220 Mathematical Proof (UBC 2023). However have exapnded the material and examples using the following books:

\begin{itemize}
	\item Main Teextbook: PLP (an introduction to mathematical proof). Link: \href{https://personal.math.ubc.ca/~PLP/}{PLP website}
	\item Book of Proof (3rd Edition) By Richard Hammack.
	\item Mathematical Proofs: A Transition to Advanced Mathematics By Chartrand et. al.
	\item Math proof lectures on YouTube: \href{https://www.youtube.com/watch?v=UuETUEJo0rk&list=PL2419488168AE7001&index=1}{YouTube Link} 
\end{itemize}



Also some useful information can be found here which are the course content of this course in previous years.
\begin{itemize}
	\item \url{https://personal.math.ubc.ca/~ilaba/teaching/math220_F2015/}
	\item \url{https://secure.math.ubc.ca/php/MathNet/courseinfo.php?session=2020W&t=outline&name=220:101}
\end{itemize}


Some open text books also can be found here in this link: \url{https://aimath.org/textbooks/approved-textbooks/} \newline


Also I will add some material from the book "A first course in logic" by Hedman.


\textbf{The rational behind mathematical proof:} The mathematics taught in the elementary courses like elementary calculus, elementary linear algebra, elementary differential equations, etc. are not good representatives of mathematics. Because those elementary courses are designed in a recipe and script oriented way to solve problems. In other words, they focus on "how" rather than "why". But in advanced mathematics the focus is on "why". That is why the mathematical logic becomes very important.  


\newpage
\tableofcontents

\newpage
\input{Sections/Fundamentals.tex}
\newpage
\input{Sections/directProof.tex}
\newpage
\input{Sections/MoreLogicAndProofByContrapositive.tex}
\newpage
\input{Sections/sets.tex}


\end{document}