\chapter{Nature of Computing}
This chapter contains the solutions to the exercises of the book ``Nature of Computing''.


\section{Chapter 1}

\begin{problem}[Hand shake theorem]
	Prove that in any finite graph, the number of verticies with odd degree is even.
\end{problem}
\begin{solution}
	Observe that the sum of degrees of all vertices is an even number,
	\[ \sum_{x\in N} \deg(x) = 2\abs{E}. \]
	So there should be even number of verticies that has odd degree (other wise the sum above will be an odd number).
\end{solution}

\begin{problem}[Pigeons and holes]
	Show that in any finite simple graph with more than one vertex, there is at least one pair of vertices that have the same degree.
\end{problem}
\begin{solution}
	We consider three exhaustive cases: The graph $ G $ has:
	\begin{enumerate}[(i)]
		\item no isolated nodes: Then the $ \deg $ function is
		\[ \deg: \set{n_1,\cdots,n_N} \to \set{1,\cdots,N-1}, \]
		This function can not be one-to-one, because the size of the range is smaller than the size of the domain. So $ \exists n_i,n_j $ such that $ \deg(n_i) = \deg(n_j) $.
		\item Exactly one isolated nodes: without loss of generality, assume $ n_1 $ is the isolated node. So $ \deg $ restricted on the non-isolated nodes will be
		\[ \deg: \set{n_2,\cdots,n_N} \to \set{1,\cdots,N-2}. \]
		For the same reason as above $ \deg $ is not one-to-one.
		
		\item More than one isolated nodes: Then immediately we have at least two nodes that both have zero degree.
	\end{enumerate}
\end{solution}

\begin{problem}[Proving Euler's Claim]
	Prove that having vertices with even degree is sufficient for a connected graph to be Eulerian. \emph{Hint: First show that if every vertex has even degree, we can cover the graph with a set of cycles such that every edge appears exactly once. Then consider combining cycles.}
\end{problem}

\begin{solution}
	First, we show that if the degree of all nodes is even, then one can cover the graph with independent cycles. Starting with a non-isolated node $ n_1 $, let $ c_1 $ be a cycle that contains $ n_1 $. To show that such a cycle exists, let $ \tilde{c}_1 $ be a $ n_1\to\tilde{n}_1 $ simple path, where $ \tilde{n}_1 $ is a neighbor of $ n_1 $ that is not on the path. Note that such a neighbor exists, because $ \deg{n_1} $ is non zero and even). Let $ c_1 $ be the path that one gets by connecting the path $ \tilde{c}_1 $ to $ n_1 $. Now delete all the edges in $ c_1 $ from the graph $ G $ to get the new graph $ G_1 $. Observe that every node in $ G_1 $ still has even degree, with possibly more isolated nodes compared to $ G $. We continue this process. Since $ G $ is finite, then this process will terminate. The set of paths $ \set{c_1,\cdots,c_n} $ are independent simple cycles.
	For two cycles $ c_i $ and $ c_j $ that share a node $ n $, we merge then according to the following diagram
	\input{Images/cycleMapping.tex} 
	By this construction $ c_{ij} $ is a bigger non-simple cycle (it is non-simple because it intersects itself). By continuing merging cycles to one another, we eventually get the Euler cycle.

\end{solution}