\begin{problem}
	The French roulette game has slots numbered from 0 to 36. The slot 0 is green, Among the slots from 1 to 36, 18 are black and 18 are red. Alex goes to a casino to play roulette. Their strategy is to always bet ``red''. They start with 50 coins, play 1 coin each turn, and stop when reaching 100 or getting broke.
	\begin{enumerate}[(a)]
		\item What is the probability that Alex reaches 100?
		\item How many coins should Alex start with to have about $50\%$ chance to reach 100?
	\end{enumerate}
\end{problem}

\begin{solution}
	\begin{enumerate}[(a)]
		\item
			Let $B$ be the event $B = \set{T_{100} < T_0}$ and we are looking for $\prob_a(B)$ where $0 \leq a \leq 100$ and indicates the number of coins we are starting with. First observe that
		\begin{itemize}
			\item $p_0 = 0$: Since if we start with zero coins we are already broken and the game is over.
			\item $p_{100} = 1$: Since if we start with 100 coins then we won the game and the game is finished.
		\end{itemize}
		To compute the probability for intermediate values of $a$, we do the first step argument. Let $WF$ be the event where Alex wins the first bet, and $LF$ the event where Alex loses the first bet. Then we can write
		\[ p_a = \prob_a(B) = \prob_a(B|WF)\prob_a(WF) + \prob_a(B|LF) \prob_a(LF). \]
		Since there are 18 red spots, then the chance to win the first bet is
		\[ \prob_a(WF) = \frac{18}{37}. \]
		and since there are 19 non-red spots in total, then the chance to win is
		\[ \prob_a(LF)  = \frac{19}{37}.\]
		Also, from Markov property, we know that
		\[ \prob_a(B|WF)= p_{a+1}, \qquad \prob_a(B|LF)=p_a. \]
		Thus the first step argument formula will be
		\[  p_a = \frac{18}{37}p_{a+1} + \frac{19}{37}p_{a-1} \implies \boxed{37p_a = 18p_{a+1} + 19p_{a-1}}. \]
		The characteristic equation for the recursive equation is
		\[ 37 = 18 x + \frac{19}{x} \implies \boxed{18x^2 - 37x + 19 = 0}. \]
		We can write it as $(x-1)(18x-19) = 0$. Thus the roots will be
		\[ r_1 = 1, \qquad r_2 = \frac{19}{18}. \]
		So
		\[ p_a = A + Br_2^a. \]
		To fine $A$ and $B$ we use the fact $p_0 = 0$, and $p_{100} = 1$. Then $A = -B$, and $A = 1/(1-r_2^{100})$. Thus
		\[ \boxed{p_a = \frac{1-r_2^a}{1-r_2^{100}}}.  \]
		
		\item  We basically need to compute find $a$ for which $p_a = 1/2$. Thus we need to solve for $a$
		\[ \frac{1-r_2^a}{1-r_2^{100}} = \frac{1}{2}. \]
		After some algebra we will find
		\[  \boxed{a = \frac{\ln(\frac{1+r_2^{100}}{2})}{\ln(r_2)} \approx 87.26}.  \]
		Thus we need to start with at least 88 coins to have a $50\%$ chance of winning.
	\end{enumerate}
	
	\qed

	
\end{solution}
